---
title: "Standardisation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE}
library(SesamThesis)
```

We read in the file pertaining to health and reproductive event data and prepare it for inclusion in the behavioural data. pivol_longer is used to merge all event names into a single "event" column and all the dates into a single column "date". For defining the "healthy" baseline period, we need to remove all dates 14 days either side of an event from the health and reproductive event data - we add these date boundaries to the data here.
```{r}
load("~/sesam/MSc_project/Final_results/Data/Health_GW50.RData")
tbl_hd_gw50 <- Health_GW50 %>%
  select(-Other_event)

### Pivot the data to merge all the event names into a single column "event" and all the dates into a single column "date"
tbl_event_gw50 <- tbl_hd_gw50 %>%
  pivot_longer(!SENSOR_MAC,
               names_to = "event",
               values_to = "date",
               values_drop_na = TRUE)
tbl_event_gw50$date <- as.Date(tbl_event_gw50$date)
head(tbl_event_gw50)

tbl_event_gw50 <- tbl_event_gw50 %>%
  mutate(pre_date = date - 14,
         post_date = date + 14)
head(tbl_event_gw50)
```

We load the behavioural dataframe created following previous data processing
```{r}
load("~/sesam/MSc_project/Final_results/Data/all_activity_SESAM_COW2.RData")
```

For standardiastaion, we first need to create a dataframe with only data pertaining to the baseline "healthy" period in which no event occurred:
```{r}
tbl_norm_data <- NULL
for (idx in 1:nrow(all_activity)){
  current_sensor <- all_activity$SENSOR_MAC[idx]
  vec_event_rows <- which(tbl_event_gw50$SENSOR_MAC == current_sensor)
  if (length(vec_event_rows) == 0) {
    tbl_norm_data <- dplyr::bind_rows(tbl_norm_data, all_activity[idx,])
    next
  }
  # if sensor ID is found in the health event data
  tbl_event_current_sensor <- tbl_event_gw50[vec_event_rows,]
  # loop over events in tbl_event_current_sensor and check whether current
  # activity is inside of an event time span
  day_current_activity <- all_activity$DAY[idx]
  b_include_rec <- TRUE
  for (jdx in 1:nrow(tbl_event_current_sensor)){
    if (day_current_activity >  tbl_event_current_sensor$pre_date[jdx] &
        day_current_activity < tbl_event_current_sensor$post_date[jdx]) {
      b_include_rec <- FALSE
    }
  }
  # check whether we have to include
  if (b_include_rec){
    tbl_norm_data <- dplyr::bind_rows(tbl_norm_data, all_activity[idx,])
  }
}

```

Calculation of the behaviour mean during the no event period:
```{r}
df_mean <- tbl_norm_data %>%
  group_by(SENSOR_MAC) %>%
  summarise(Lying = mean(Lying), Eating = mean(Eating), Standing = mean(Standing), Walking = mean(Walking), Ruminating = mean(Ruminating), Activity = mean(Daily_Activ))

mean_merge <- melt(subset(df_mean,
                          select=c("SENSOR_MAC",
                                   "Lying",
                                   "Eating",
                                   "Standing",
                                   "Walking",
                                   "Ruminating")),
                   id.vars = c("SENSOR_MAC"),
                   variable.name = 'activity')
colnames(mean_merge) <- c("SENSOR_MAC", "series", "mean_minutes")
```

Calculation of the behaviour standard deviation during the no event period 
```{r}
df_SD <- tbl_norm_data %>%
  group_by(SENSOR_MAC) %>%
  summarise(Lying = sd(Lying), Eating = sd(Eating), Standing = sd(Standing), Walking = sd(Walking), Ruminating = sd(Ruminating), Activity = sd(Daily_Activ))

SD_merge <- melt(subset(df_SD,
                        select=c("SENSOR_MAC",
                                 "Lying",
                                 "Eating",
                                 "Standing",
                                 "Walking",
                                 "Ruminating")),
                 id.vars = c("SENSOR_MAC"),
                 variable.name = 'activity')
colnames(SD_merge) <- c("SENSOR_MAC", "series", "SD_minutes")
```

Melt all columns together into the behaviour dataset pre-processed to melt behaviours into a single column. Then we add a new column for original behaviour value - mean.
Standardise by dividing this new column by the SD.
```{r}
load("~/sesam/MSc_project/GW50/update/data/melted_SESAM_COW.RData")
df2_with_mean <- merge(df2, mean_merge, by = c("SENSOR_MAC", "series"))
df2_mean_SD <- merge(df2_with_mean, SD_merge, by = c("SENSOR_MAC", "series"))
df_normalisation <- df2_mean_SD

final_df_normalisation <- df_normalisation %>%
  mutate(new_value = value - as.numeric(mean_minutes))

### standardisation equation: ((value-mean of no event data)/sd of no event data). We multiply by 10 to scale.
standardised_df <- final_df_normalisation %>%
  mutate(standarized_value = (new_value/as.numeric(SD_minutes)))

save(standardised_df,
     file = "~/sesam/MSc_project/Final_results/Data/standardised_df.RData",
     verbose=T)
```

