---
title: "SVM_calving"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Load packages
```{r}
library(caret)
library(dplyr)
library(ggplot2)
library(DMwR)
library(ROSE)
```

Load and filter data
```{r}
load("~/sesam/MSc_project/Final_results/Data/df_prep50_health.RData")
df_prep <- df_prep50 %>%
  filter(Event == "no_event" | Event == "Pre_calving")

df_prep <- df_prep %>%
  select(-SENSOR_MAC, - DAY)


df_prep$Event <- as.factor(df_prep$Event)
table(df_prep$Event)
```

#Training control: cross validation, 3 repeats
```{r}
trainSetControl <- trainControl(method = "repeatedcv",
                                number = 10,
                                repeats = 3,
                                search = "grid")
```

# 80:20 dataset
## Test/train split
```{r}
set.seed(100)
indexes <- createDataPartition(df_prep$Event,
                               times = 1,
                               p = 0.8,
                               list = FALSE)

trainingSet <- df_prep[indexes,]
testingSet <- df_prep[-indexes,]

trainingSet <- na.omit(trainingSet)
testingSet <- na.omit(testingSet)

table(trainingSet$Event)
table(testingSet$Event)
```

## Data sampling for training data
```{r}
#Random under sampling
rus_train  <- downSample(x = trainingSet[,-1],
                                   y = trainingSet$Event,
                                   yname = "Event")
table(rus_train$Event)
save(rus_train, file = "~/sesam/MSc_project/Final_results/Data/SVM/rus_train.RData")

#Random over sampling
ros_train <- upSample(x = trainingSet[,-1],
                    y = trainingSet$Event,
                    yname = "Event")
table(ros_train$Event)
save(ros_train, file = "~/sesam/MSc_project/Final_results/Data/SVM/ros_train.RData")


#SMOTE
str(trainingSet)
trainingSet$Event <- as.factor(trainingSet$Event)
trainingSet <- as.data.frame(trainingSet)
SMOTE_train <- SMOTE(Event ~., trainingSet, perc.over = 1000)
table(SMOTE_train$Event)
save(SMOTE_train, file = "~/sesam/MSc_project/Final_results/Data/SVM/SMOTE_train.RData")


#ROSE
ROSE_train <- ROSE(Event ~ ., data = trainingSet)$data
table(ROSE_train$Event)
save(ROSE_train, file = "~/sesam/MSc_project/Final_results/Data/SVM/ROSE_train.RData")

```

##Linear SVM 
###Tuning grid for selecting C parameter
```{r}
tuningGrid <- expand.grid(C = seq(0.1,2, length = 20))
```

###Model training 
```{r}
set.seed(100)
SVM_unbalanced_lin <- train(Event ~ ., data = trainingSet,
                  method = "svmLinear",
                  tuneGrid = tuningGrid,
                  trControl = trainSetControl)
saveRDS(SVM_unbalanced_lin, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVMlin_80_unbalanced.rds")


### random under sampling
set.seed(100)
SVM_rus_lin <- train(Event ~ ., data = rus_train,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_rus_lin, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVMlin_80_rus.rds")


### random over sampling
set.seed(100)
SVM_ros_lin <- train(Event ~ ., data = ros_train,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_ros_lin, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVMlin_80_ros.rds")


### SMOTE
set.seed(100)
SVM_SMOTE_lin <- train(Event ~ ., data = SMOTE_train,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_SMOTE_lin, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVMlin_80_SMOTE.rds")
### ROSE
set.seed(100)
SVM_ROSE_lin <- train(Event ~ ., data = ROSE_train,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_ROSE_lin, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVMlin_80_ROSE.rds")
```

###Predictions using testing data 
```{r}
pred_SVM_unbalanced <- predict(SVM_unbalanced_lin, testingSet)
predict_SVM_rus <- predict(SVM_rus_lin, testingSet)
predict_SVM_ros <- predict(SVM_ros_lin, testingSet)
predict_SVM_SMOTE <- predict(SVM_SMOTE_lin, testingSet)
predict_SVM_ROSE <- predict(SVM_ROSE_lin, testingSet)
```

###Confusion matrix 
```{r}
confusionSVM_unbalanced <- confusionMatrix(pred_SVM_unbalanced, positive = "Pre_calving", mode = "everything", testingSet$Event)
confusionSVM_unbalanced
confusionSVM_rus <- confusionMatrix(predict_SVM_rus, positive = "Pre_calving", mode = "everything", testingSet$Event)
confusionSVM_rus
confusionSVM_ros <- confusionMatrix(predict_SVM_ros, positive = "Pre_calving", mode = "everything", testingSet$Event)
confusionSVM_ros
confusionSVM_SMOTE <- confusionMatrix(predict_SVM_SMOTE, positive = "Pre_calving", mode = "everything", testingSet$Event)
confusionSVM_SMOTE
confusionSVM_ROSE <- confusionMatrix(predict_SVM_ROSE, positive = "Pre_calving", mode = "everything", testingSet$Event)
confusionSVM_ROSE
```

#70:30 data split 
## Train/test data 
```{r}
set.seed(100)
indexes2 <- createDataPartition(df_prep$Event,
                                times = 1,
                                p = 0.7,
                                list = FALSE)

trainingSet2 <- df_prep[indexes2,]
testingSet2 <- df_prep[-indexes2,]

trainingSet2 <- na.omit(trainingSet2)
testingSet2 <- na.omit(testingSet2)

table(trainingSet2$Event)
table(testingSet2$Event)
```

##Data sampling
```{r}
### Random under sampling
set.seed(100)
rus_train2  <- downSample(x = trainingSet2[,-1],
                         y = trainingSet2$Event,
                         yname = "Event")
table(rus_train2$Event)
save(rus_train2, file = "~/sesam/MSc_project/Final_results/Data/SVM/rus_train2.RData")

### Random over sampling
set.seed(100)
ros_train2 <- upSample(x = trainingSet2[,-1],
                      y = trainingSet2$Event,
                      yname = "Event")
table(ros_train2$Event)
save(ros_train2, file = "~/sesam/MSc_project/Final_results/Data/SVM/ros_train2.RData")

### SMOTE
set.seed(100)
str(trainingSet2)
trainingSet2$Event <- as.factor(trainingSet2$Event)
trainingSet2 <- as.data.frame(trainingSet2)
SMOTE_train2 <- SMOTE(Event ~., trainingSet2, perc.over = 1000)
table(SMOTE_train2$Event)
save(SMOTE_train2, file = "~/sesam/MSc_project/Final_results/Data/SVM/SMOTE_train2.RData")

### ROSE
set.seed(100)
ROSE_train2 <- ROSE(Event ~ ., data = trainingSet2)$data
table(ROSE_train2$Event)
save(ROSE_train2, file = "~/sesam/MSc_project/Final_results/Data/SVM/ROSE_train2.RData")
```

##SVM 
### Model training 
```{r}
### unbalanced
set.seed(100)
SVM_unbalanced_lin_70 <- train(Event ~ ., data = trainingSet2,
                        method = "svmLinear",
                        tuneGrid = tuningGrid,
                        trControl = trainSetControl)
saveRDS(SVM_unbalanced_lin_70, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_unbalanced_lin_70.rds")

### random under sampling
set.seed(100)
SVM_rus_lin_70 <- train(Event ~ ., data = rus_train2,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)

saveRDS(SVM_rus_lin_70, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_rus_lin_70.rds")


### random over sampling
set.seed(100)
SVM_ros_lin_70 <- train(Event ~ ., data = ros_train2,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)

saveRDS(SVM_ros_lin_70, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_ros_lin_70.rds")


### SMOTE
set.seed(100)
SVM_SMOTE_lin_70 <- train(Event ~ ., data = SMOTE_train2,
                   method = "svmLinear",
                   tuneGrid = tuningGrid,
                   trControl = trainSetControl)
saveRDS(SVM_SMOTE_lin_70, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_SMOTE_lin_70.rds")


### ROSE
set.seed(100)
SVM_ROSE_lin_70 <- train(Event ~ ., data = ROSE_train2,
                  method = "svmLinear",
                  tuneGrid = tuningGrid,
                  trControl = trainSetControl)
saveRDS(SVM_ROSE_lin_70, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_ROSE_lin_70.rds")
```

### Predictions of test data 
```{r}
pred_SVM_unbalanced_lin_70 <- predict(SVM_unbalanced_lin_70, testingSet2)
predict_SVM_rus_lin_70 <- predict(SVM_rus_lin_70, testingSet2)
predict_SVM_ros_lin_70 <- predict(SVM_ros_lin_70, testingSet2)
predict_SVM_SMOTE_lin_70 <- predict(SVM_SMOTE_lin_70, testingSet2)
predict_SVM_ROSE_lin_70 <- predict(SVM_ROSE_lin_70, testingSet2)
```

### Confusion matrix
```{r}
confusionSVM_unbalanced_lin_70 <- confusionMatrix(pred_SVM_unbalanced_lin_70, positive = "Pre_calving", mode = "everything", testingSet2$Event)
confusionSVM_unbalanced_lin_70
confusionSVM_rus_lin_70 <- confusionMatrix(predict_SVM_rus_lin_70, positive = "Pre_calving", mode = "everything", testingSet2$Event)
confusionSVM_rus_lin_70
confusionSVM_ros_lin_70 <- confusionMatrix(predict_SVM_ros_lin_70, positive = "Pre_calving", mode = "everything", testingSet2$Event)
confusionSVM_ros_lin_70
confusionSVM_SMOTE_lin_70 <- confusionMatrix(predict_SVM_SMOTE_lin_70, positive = "Pre_calving", mode = "everything", testingSet2$Event)
confusionSVM_SMOTE_lin_70
confusionSVM_ROSE_lin_70 <- confusionMatrix(predict_SVM_ROSE_lin_70, positive = "Pre_calving", mode = "everything", testingSet2$Event)
confusionSVM_ROSE_lin_70
```

# 60:40 data split
## train/test data 
```{r}
set.seed(100)
indexes3 <- createDataPartition(df_prep$Event,
                                times = 1,
                                p = 0.6,
                                list = FALSE)

trainingSet3 <- df_prep[indexes3,]
testingSet3 <- df_prep[-indexes3,]

trainingSet3 <- na.omit(trainingSet3)
testingSet3 <- na.omit(testingSet3)

table(trainingSet3$Event)
table(testingSet3$Event)
```

## data sampling
```{r}
### Random under sampling
rus_train3  <- downSample(x = trainingSet3[,-1],
                         y = trainingSet3$Event,
                         yname = "Event")
table(rus_train3$Event)
save(rus_train3, file = "~/sesam/MSc_project/Final_results/Data/SVM/rus_train3.RData")

### Random over sampling
ros_train3 <- upSample(x = trainingSet3[,-1],
                      y = trainingSet3$Event,
                      yname = "Event")
table(ros_train3$Event)
save(ros_train3, file = "~/sesam/MSc_project/Final_results/Data/SVM/ros_train3.RData")

### SMOTE
str(trainingSet3)
trainingSet3$Event <- as.factor(trainingSet3$Event)
trainingSet3 <- as.data.frame(trainingSet3)
smote_train3 <- SMOTE(Event ~., trainingSet3, perc.over = 1000)
table(smote_train3$Event)
save(smote_train3, file = "~/sesam/MSc_project/Final_results/Data/SVM/smote_train3.RData")
table(smote_train3)

### ROSE
rose_train3 <- ROSE(Event ~ ., data = trainingSet3)$data
table(rose_train3$Event)
save(rose_train3, file = "~/sesam/MSc_project/Final_results/Data/SVM/rose_train3.RData")
```

##SVM
### Model training
```{r}
### unbalanced
set.seed(100)
SVM_unbalanced_lin60 <- train(Event ~ ., data = trainingSet3,
                        method = "svmLinear",
                        tuneGrid = tuningGrid,
                        trControl = trainSetControl)
saveRDS(SVM_unbalanced_lin60, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_unbalanced_lin60.rds")


### random under sampling
set.seed(100)
SVM_rus_lin60 <- train(Event ~ ., data = rus_train3,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_rus_lin60, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_rus_lin60.rds")


### random over sampling
set.seed(100)
SVM_ros_lin60 <- train(Event ~ ., data = ros_train3,
                 method = "svmLinear",
                 tuneGrid = tuningGrid,
                 trControl = trainSetControl)
saveRDS(SVM_ros_lin60, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_ros_lin60.rds")


### SMOTE
set.seed(100)
SVM_SMOTE_lin60 <- train(Event ~ ., data = smote_train3,
                   method = "svmLinear",
                   tuneGrid = tuningGrid,
                   trControl = trainSetControl)
saveRDS(SVM_SMOTE_lin60, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_SMOTE_lin60.rds")


### ROSE
set.seed(100)
SVM_ROSE_lin60 <- train(Event ~ ., data = rose_train3,
                  method = "svmLinear",
                  tuneGrid = tuningGrid,
                  trControl = trainSetControl)
saveRDS(SVM_ROSE_lin60, file = "~/sesam/MSc_project/Final_results/Data/SVM/SVM_ROSE_lin60.rds")
```

### Predictions of test data 
```{r}
pred_SVM_unbalanced_lin60 <- predict(SVM_unbalanced_lin60, testingSet3)
predict_SVM_rus_lin60 <- predict(SVM_rus_lin60, testingSet3)
predict_SVM_ros_lin60 <- predict(SVM_ros_lin60, testingSet3)
predict_SVM_SMOTE_lin60 <- predict(SVM_SMOTE_lin60, testingSet3)
predict_SVM_ROSE_lin60 <- predict(SVM_ROSE_lin60, testingSet3)
```

### Confusion matrix 
```{r}
confusionSVM_unbalanced_lin60 <- confusionMatrix(pred_SVM_unbalanced_lin60, positive = "Pre_calving", mode = "everything", testingSet3$Event)
confusionSVM_unbalanced_lin60
confusionSVM_rus_lin60 <- confusionMatrix(predict_SVM_rus_lin60, positive = "Pre_calving", mode = "everything", testingSet3$Event)
confusionSVM_rus_lin60
confusionSVM_ros_lin60 <- confusionMatrix(predict_SVM_ros_lin60, positive = "Pre_calving", mode = "everything", testingSet3$Event)
confusionSVM_ros_lin60
confusionSVM_SMOTE_lin60 <- confusionMatrix(predict_SVM_SMOTE_lin60, positive = "Pre_calving", mode = "everything", testingSet3$Event)
confusionSVM_SMOTE_lin60
confusionSVM_ROSE_lin60 <- confusionMatrix(predict_SVM_ROSE_lin60, positive = "Pre_calving", mode = "everything", testingSet3$Event)
confusionSVM_ROSE_lin60
```

